Assessing Gender Bias in Machine Translation -- A Case Study with Google Translate

### 2018-03-13 Paper/main.tex
AA

On a 2014 article, Londa Schiebinger suggested that scientific research fails to take gender issues into account, arguing that the phenomenon of male defaults on new technologies such as Google Translate provides a window into this asymmetry \cite{schiebinger2014scientific}. Since then, recent worrisome results in machine learning have somewhat supported Schiebinger's view, and perhaps even partially confirmed some of Chomsky's fears. Not only Google photos' statistical image labeling algorithm has been found to classify dark-skinned people as gorillas \cite{garcia2016racist} and purportedly intelligent programs have been suggested to be negatively biased against black prisoners when predicting criminal behavior \cite{angwin2016machine} but the machine learning revolution has also indirectly revived heated debates about the controversial field of physiognomy, with proposals of AI systems capable of identifying the sexual orientation of an individual through its facial characteristics \cite{wang2017deep}. Similar concerns are growing at an unprecedented rate in the media, with reports of Apple's Iphone X face unlock feature failing to differentiate between two different Asian people \cite{womanunlockphone2017} and automatic soap dispensers which reportedly do not recognize black hands \cite{racistsoapdispenser2017}. \emph{Machine bias}, the phenomenon by which trained statistical models unbeknownst to their creators become vessels of their own prejudices, is growing into a pressing concern for the modern times, and is an invitation for us to ask ourselves whether there are limits to our dependence on these techniques -- and more importantly, whether some of these limits have already been traversed.

With this in mind, we propose a quantitative analysis of the phenomenon of gender bias in machine translation. We illustrate how this can be done by simply exploiting Google Translate to map sentences from a gender neutral language into English. As Figure \ref{fig:screenshot-gtranslate-hungarian} exemplifies, this approach produces results consistent with the hypothesis that senteces about stereotypical gender roles are translated accordingly with high probability: \emph{nurse} and \emph{baker} are translated with female pronouns while \emph{engineer} and \emph{CEO} are translated with male ones.

\begin{figure}[h]
	\centering
	\fbox{\includegraphics[width=14cm]{pictures/screenshot-gtranslate-hungarian}}
	\caption{Translating sentences from a gender neutral language such as Hungarian to English provides a glimpse into the phenomenon of gender bias in machine translation. This screenshot from Google Translate shows how occupations from traditionally male-dominated fields such as scholar, engineer and CEO are interpreted as male, while occupations such as nurse, baker and wedding organizer are interpreted as female.}
	\label{fig:screenshot-gtranslate-hungarian}
\end{figure}
If automatic translation tools are indeed gender biased as Figure \ref{fig:screenshot-gtranslate-hungarian} seems to suggest, then possibly this phenomenon can be probed to yield many valuable insights. One could in principle group different professional occupations according to their broad category (artistic, scientific, industrial, etc.) and evaluate whether the proportion of male pronouns is more pronounced in some fields compared to others. We expect, for example, that most sentences about computer-related jobs will be translated with male pronouns, as a reflection of the gender asymmetry of this job market. We can also collect translation statistics among varied gender neutral languages to see whether different cultures make different assumptions about the gender of a subject.

If conclusive, in our view such an analysis would suggest that automatic translation tools should be designed in such a way as to avoid becoming hostages to their own underlying statistical models. Because these tools are trained with real world data they implicitly absorb the biases and stereotypes of our society, which must then be painstakingly removed. This poses 
mostly an ethical rather than technical limitation to statistical machine translation and big data statistical models in general.

@@

@@

\begin{table}[H]
\small{
	\centering
	\begin{tabular}{|c|c|c|c|c|}
	\hline
	stagehands & author & neurologist \\ \hline
	screenwriter & animator & marketing director \\ \hline
	biochemist & endocrinologist & freelancer \\ \hline
	neurosurgeon & computer scientist & petrochemical engineer \\ \hline
	food stylist & cardiothoracic surgeon & property master \\ \hline
	literary editor & video editor & animation director \\ \hline
	house manager & chief administrative officer & arts administration \\ \hline
	actor & dialysis technician & family nurse practitioner \\ \hline
	psychologist & chief creative officer & flash developer \\ \hline
	scenic artist & producer & medical laboratory scientist \\ \hline
	\end{tabular}
	\caption{A randomly selected example subset of thirty occupations obtained from our dataset with a total of 462 different occupations.}
	\label{tab:occupations-examples}
	}
\end{table}

@@

\begin{table}[H]
\small{
	\centering
	\begin{tabular}{|c|c|c|c|c|c|}
	\hline
	Category & Female & Male & Neutral & Ratio & Total \\ \hline
	\hline
	Artistic        & 188 & 518 & 72 & 2.755 	& 918 \\ \hline
	Computer        & 7   & 125 & 12 & 17.857 	& 171 \\ \hline
	Corporate       & 36  & 340 & 23 & 9.444 	& 450 \\ \hline
	Dance           & 31  & 33  & 8  & 1.064 	& 81  \\ \hline
	Film-television & 54  & 125 & 18 & 2.315 	& 234 \\ \hline
	Healthcare      & 176 & 483 & 64 & 2.744 	& 792 \\ \hline
	Industrial      & 40  & 135 & 29 & 3.375   	& 234 \\ \hline
	Science         & 34  & 357 & 25 & 10.500	& 450 \\ \hline
	Service         & 5   & 63  & 10 & 12.600	& 90  \\ \hline
	Theatre         & 75  & 296 & 38 & 3.947 	& 477 \\ \hline
	Writing         & 41  & 148 & 30 & 3.610 	& 261 \\ \hline
	\hline
	Total           & 687 & 2623 & 329 & 3.818 & 4158 \\ \hline
	\end{tabular}
	\caption{Number of female, male and neutral pronominal genders per occupation category in the translated sentences. The corresponding sex ratios (\# Male / \# Female) show just how much male defaults are prominent in male dominated fields such as computer science, with up to $\approx 18$ occurrences of male pronouns for each of a female one.}
	\label{tab:gender-by-category}
	}
\end{table}

\begin{figure}[H]
	\centering
	\includegraphics[width=10cm]{pictures/gender-by-category}
	\caption{The distribution of pronominal genders in the translated sentences not only suggests a tendency towards male defaults but also reflects the structure of male dominated fields, with the proportion of male pronouns amounting to $73\%$ in computer related jobs and $76\%$ in corporate jobs respectively. Because Google Translate occasionally fails to translate a sentence, the bars for some categories fail to add up to $100\%$.}
	\label{fig:gender-by-category}
\end{figure}

@@

\begin{table}[H]
\small{
	\centering
	\begin{tabular}{|c|c|c|c|c|c|}
	\hline
	Language & Female & Male & Neutral & Ratio & Total \\ \hline
	\hline
	Malay     & 47  & 415  & 0   & 8.830 	& 462 \\ \hline
	Estonian  & 130 & 332  & 0   & 2.554 	& 462 \\ \hline
	Finnish   & 179 & 283  & 0   & 1.581 	& 462 \\ \hline
	Hungarian & 189 & 270  & 1   & 1.429 	& 462 \\ \hline
	Armenian  & 101 & 360  & 1   & 3.564 	& 462 \\ \hline
	Japanese  & 2   & 215  & 0   & 107.500  & 462 \\ \hline
	Turkish   & 22  & 394  & 43  & 17.909	& 462 \\ \hline
	Yoruba    & 12  & 304  & 131 & 25.333	& 462 \\ \hline
	Basque    & 5   & 50   & 153 & 10.000   & 462 \\ \hline
	Swahili   & 76  & 386  & 0   & 5.079 	& 462 \\ \hline
	Chinese   & 14  & 385  & 23  & 27.500   & 462 \\ \hline
	\hline
	Total     & 777 & 3394 & 352 & 4.368	& 4158   \\ \hline
	\end{tabular}
	\caption{Number of female, male and neutral pronominal genders per language in the translated sentences. The corresponding sex ratios (\# Male / \# Female) show just how much male defaults are prominent in some languages such as Chinese, with almost 30 male pronouns for each female one.}
	\label{tab:gender-by-language}
	}
\end{table}

@@

Instead of grouping sentences either by category or language, we can also visualize each of them individually on a scatter plot. Figure \ref{fig:scatterplot-languages} shows each occupation as a point on a bi-dimensional lattice, arranged horizontally by their sex ratio and vertically by the proportion of gender neutral pronouns, both averaged over translations from all tested languages. Each point is also color coded according to that occupation's respective category (Artistic, Writing, Science, etc.).

\begin{figure}[H]
	\centering
	\includegraphics[width=10cm]{pictures/scatterplot-languages}
	\caption{Scatter plot of the translated sentences' statistics. Each point, corresponding to a single occupation, is positioned in the horizontal axis according to its sex ratio and in the vertical axis according to the frequency with which it is translated using gender neutral pronouns. Both quantities are averaged over all tested languages.}
	\label{fig:scatterplot-languages}
\end{figure}

@@

@@

\section{Conclusions}
In this paper, we have provided evidence that statistical translation tools such as Google Translate exhibit implicit gender biases and a tendency towards male defaults, which possibly stem from the real world data used to train them. As a result, we suggest that such tools can be probed to yield insights about stereotypical gender roles in our society. By translating sentences from gender neutral languages such as Hungarian and Chinese into English, we were able to collect statistics about the asymmetry between female and male pronominal genders in the translation process. Because Google Translate typically uses English as a \emph{lingua franca} to translate between other languages, our findings probably generalize to most translations from gender neutral idioms, although we have not tested this hypothesis yet. Our results further show that although male pronouns are typical, the proportion of female pronouns can vary significantly according either to the adjectives used (\say{shy}/\say{happy} vs \say{brave}/\say{guilty}) or the category of professional occupations (artistic jobs are far more likely to be translated with female pronouns than computer-related ones). We have also shown that different languages are differently biased towards male defaults, with Hungarian exhibiting a better balance between male and female pronouns than, say, Chinese. Some languages such as Yoruba and Basque have been found to translate sentences with gender neutral pronouns very often, although this is the exception rather than the rule.

We think this work can shed further light on some of the technical and ethical difficulties that arise from statistical machine translation, and hope that it will lead to discussions about the role of AI engineers on minimizing potentially harmful effects of the pressing modern concerns of machine bias.

\begin{thebibliography}{xx}

@@

@@

@@

@@

@@

@@

\end{thebibliography}
\end{document}

>>>>

On a 2014 article, Londa Schiebinger suggested that scientific research fails to take gender issues into account, arguing that the phenomenon of male defaults on new technologies such as Google Translate provides a window into this asymmetry \cite{schiebinger2014scientific}. Since then, recent worrisome results in machine learning have somewhat supported Schiebinger's view. Not only Google photos' statistical image labeling algorithm has been found to classify dark-skinned people as gorillas \cite{garcia2016racist} and purportedly intelligent programs have been suggested to be negatively biased against black prisoners when predicting criminal behavior \cite{angwin2016machine} but the machine learning revolution has also indirectly revived heated debates about the controversial field of physiognomy, with proposals of AI systems capable of identifying the sexual orientation of an individual through its facial characteristics \cite{wang2017deep}. Similar concerns are growing at an unprecedented rate in the media, with reports of Apple's Iphone X face unlock feature failing to differentiate between two different Asian people \cite{womanunlockphone2017} and automatic soap dispensers which reportedly do not recognize black hands \cite{racistsoapdispenser2017}. \emph{Machine bias}, the phenomenon by which trained statistical models unbeknownst to their creators become vessels of their own prejudices, is growing into a pressing concern for the modern times, and is an invitation for us to ask ourselves whether there are limits to our dependence on these techniques -- and more importantly, whether some of these limits have already been traversed. In the wave of algorithmic bias, some have argued for the creation of some kind of agency in the likes of the Food and Drug Administration with the sole purpose of regulating algorithmic discrimination \cite{kirkpatrick2016battling}.

With this in mind, we propose a quantitative analysis of the phenomenon of gender bias in machine translation. We illustrate how this can be done by simply exploiting Google Translate to map sentences from a gender neutral language into English. As Figure \ref{fig:screenshot-gtranslate-hungarian} exemplifies, this approach produces results consistent with the hypothesis that sentences about stereotypical gender roles are translated accordingly with high probability: \emph{nurse} and \emph{baker} are translated with female pronouns while \emph{engineer} and \emph{CEO} are translated with male ones.

\begin{figure}[h]
	\centering
	\fbox{\includegraphics[width=14cm]{pictures/screenshot-gtranslate-hungarian}}
	\caption{Translating sentences from a gender neutral language such as Hungarian to English provides a glimpse into the phenomenon of gender bias in machine translation. This screenshot from Google Translate shows how occupations from traditionally male-dominated fields such as scholar, engineer and CEO are interpreted as male, while occupations such as nurse, baker and wedding organizer are interpreted as female.}
	\label{fig:screenshot-gtranslate-hungarian}
\end{figure}

\section{Motivation}

As of 2017, Google translate is one of the largest publicly available machine translation tools in existence, amounting 200 million users daily \cite{Gtranslate200daily2017}. Initially relying on United Nations and European Parliament transcripts to gather data, since 2014 Google Translate has inputed content from its users through the Translate Community initiative \cite{TranslateCommunity}. Recently however there has been a growing concern about gender asymmetries in the translation mechanism, with some heralding it as \say{sexist} \cite{AlgorithmGtranslateSexist2018}. This concern has to at least some extent a scientific backup: A recent study has shown that word embeddings are particularly prone to yielding gender stereotypes \cite{bolukbasi2016man}. Fortunately, the researchers propose a relatively simple \emph{debiasing} algorithm with promising results: they were able to cut the proportion of stereotypical analogies from $19\%$ to $6\%$ without any significant compromise in the performance of the word embedding procedure. The researchers are not alone: there is a growing effort to systematically discover and resolve issues of algorithmic bias in black-box algorithms \cite{hajian2016algorithmic} The success of these results suggest that a similar technique could be used to remove gender bias from Google Translate outputs, should it exist. This paper intends to investigate whether it does. We are optimistic that our research endeavors, if conclusive, can be used to argue that there is a positive payoff in redesigning modern statistical translation tools.

@@

\begin{table}[H]
\begin{small}
	\centering
	\begin{tabular}{|m{3.5cm}|c|c|c|}
	\hline	
	Category 										& Group 						& \# Occupations 	& Female Participation	\\ \hline \hline
	Education, training, and library 				& Education						& $22$ 				& $73.0\%$				\\ \hline
	Business and financial operations				& Corporate						& $46$ 				& $54.0\%$				\\ \hline
	Office and administrative support				& Service						& $87$ 				& $72.2\%$				\\ \hline
	Healthcare support 								& Healthcare					& $16$ 				& $87.1\%$				\\ \hline
	Management 										& Corporate						& $46$ 				& $39.8\%$				\\ \hline
	Installation, maintenance, and repair 			& Service						& $91$ 				& $4.0\%$ 				\\ \hline
	Healthcare practitioners and technical 			& Healthcare					& $43$ 				& $75.0\%$				\\ \hline
	Community and social service 					& Service						& $14$ 				& $66.1\%$				\\ \hline
	Sales and related 								& Corporate						& $28$ 				& $49.1\%$				\\ \hline
	Production 										& Production					& $264$ 			& $28.9\%$				\\ \hline
	Architecture and engineering 					& STEM 							& $29$ 				& $16.2\%$				\\ \hline
	Life, physical, and social science				& STEM 							& $34$ 				& $47.4\%$				\\ \hline
	Transportation and material moving				& Service						& $70$ 				& $17.3\%$				\\ \hline
	Arts, design, entertainment, sports, and media 	& Arts / Entertainment 			& $37$ 				& $46.9\%$				\\ \hline
	Legal											& Legal 						& $7$				& $52.8\%$				\\ \hline
	Protective Service 								& Service 						& $28$ 				& $22.3\%$				\\ \hline
	Food preparation and serving related 			& Service 						& $17$ 				& $53.8\%$				\\ \hline
	Farming, fishing, and forestry 					& Farming / Fishing / Forestry 	& $13$ 				& $23.4\%$				\\ \hline
	Computer and mathematical						& STEM 							& $16$ 				& $25.5\%$				\\ \hline
	Personal care and service						& Service 						& $33$ 				& $76.1\%$				\\ \hline
	Construction and extraction 					& Construction / Extraction 	& $68$ 				& $3.0\%$ 				\\ \hline
	Building and grounds cleaning and maintenance 	& Service 						& $10$ 				& $40.7\%$				\\ \hline \hline
	Total											& - 							& $1019$			& $41.3\%$				\\ \hline
	\end{tabular}
	\caption{Selected occupations obtained from the U.S. Bureau of Labor Statistics \url{https://???}, grouped by category. We obtained a total of $1019$ occupations from $22$ distinct categories. We have further grouped them into broader groups (or \emph{super-categories}) to ease analysis and visualization.}
	\label{tab:occupations}
	\end{small}
\end{table}

@@

\begin{table}[H]
\small{
	\centering
	\begin{tabular}{|c|c|c|c|c|}
	\hline
	Insurance sales agent & Editor & Rancher \\ \hline
	Ticket taker & Pile-driver operator & Tool maker \\ \hline
	Jeweler & Judicial law clerk & Auditing clerk \\ \hline
	Physician & Embalmer & Door-to-door salesperson \\ \hline
	Packer & Bookkeeping clerk & Community health worker \\ \hline
	Sales worker & Floor finisher & Social science technician \\ \hline
	Probation officer & Paper goods machine setter & Heating installer \\ \hline
	Animal breeder & Instructor & Teacher assistant \\ \hline
	Statistical assistant & Shipping clerk & Trapper \\ \hline
	Pharmacy aide & Sewing machine operator & Service unit operator \\ \hline
	\end{tabular}
	\caption{A randomly selected example subset of thirty occupations obtained from our dataset with a total of $1019$ different occupations.}
	\label{tab:occupations-examples}
	}
\end{table}

@@

\begin{table}[H]
\small{
	\centering
	\begin{tabular}{|c|c|c|c|}
	\hline
	Category 								& Female ($\%$)	& Male ($\%$)	& Neutral ($\%$) \\ \hline
	\hline		
	Education / Training / Library 			& $25.207$ 		& $59.917$		& $2.893$		\\ \hline
	Business / Financial 					& $6.324$ 		& $67.194$		& $7.312$		\\ \hline
	Office / Administrative					& $10.031$		& $62.382$		& $7.732$		\\ \hline
	Healthcare support						& $29.545$		& $51.705$		& $7.955$		\\ \hline
	Management								& $9.486$		& $68.182$		& $3.953$		\\ \hline
	Installation / Maintenance / Repair		& $3.596$		& $68.332$		& $11.588$		\\ \hline
	Healthcare practitioners / technical	& $21.564$		& $61.734$		& $6.977$		\\ \hline
	Community and social service			& $15.584$ 		& $63.636$		& $4.545$		\\ \hline
	Sales and related						& $8.117$ 		& $69.156$		& $8.766$		\\ \hline
	Production								& $14.36$ 		& $53.306$		& $13.12$		\\ \hline
	Architecture and engineering			& $2.194$		& $79.624$		& $5.016$		\\ \hline
	Life, physical, and social sciences		& $3.743$ 		& $82.888$		& $3.743$		\\ \hline
	Transportation / material moving		& $7.532$ 		& $66.234$		& $11.688$		\\ \hline
	Arts / Entertainment					& $7.862$ 		& $71.007$		& $4.423$		\\ \hline
	Legal									& $11.688$ 		& $67.532$		& $5.195$		\\ \hline
	Protective service						& $6.818$ 		& $75.0$		& $5.844$		\\ \hline
	Food prepaparation / serving			& $16.578$ 		& $59.893$		& $10.16$		\\ \hline
	Farming, fishing, and forestry			& $10.49$ 		& $62.238$		& $8.392$		\\ \hline
	Computer and mathematical				& $2.841$ 		& $76.136$		& $4.545$		\\ \hline
	Personal care and service				& $20.937$ 		& $55.096$		& $9.091$		\\ \hline
	Construction and extraction				& $9.492$ 		& $61.631$		& $11.096$		\\ \hline
	Cleaning / Maintenance					& $11.818$		& $63.636$		& $8.182$		\\ \hline
	\hline
	Total           & ? & ? & ? \\ \hline
	\end{tabular}
	\caption{Number of female, male and neutral pronominal genders per occupation category in the translated sentences. The corresponding sex ratios (\# Male / \# Female) show just how much male defaults are prominent in male dominated fields such as computer science, with up to $\approx 18$ occurrences of male pronouns for each of a female one.}
	\label{tab:gender-by-category}
	}
\end{table}

\begin{figure}[H]
	\centering
	\includegraphics[width=10cm]{pictures/gender-by-category-grouped}
	\caption{The distribution of pronominal genders in the translated sentences not only suggests a tendency towards male defaults but also reflects the structure of male dominated fields, with the proportion of male pronouns amounting to $73\%$ in computer related jobs and $76\%$ in corporate jobs respectively. Because Google Translate occasionally fails to translate a sentence, the bars for some categories fail to add up to $100\%$.}
	\label{fig:gender-by-category}
\end{figure}

@@

\begin{table}[H]
\small{
	\centering
	\begin{tabular}{|c|c|c|c|}
	\hline
	Language 	& Female ($\%$) 	& Male ($\%$)		& Neutral ($\%$)	\\ \hline
	\hline
	Malay     	& $4.122$ 			& $95.780$  		& $0.000$			\\ \hline
	Estonian  	& $18.351$			& $80.962$  		& $0.491$			\\ \hline
	Finnish   	& $33.562$			& $66.143$  		& $0.000$			\\ \hline
	Hungarian 	& $36.016$			& $63.297$  		& $0.393$			\\ \hline
	Armenian  	& $9.814$			& $81.256$  		& $0.589$			\\ \hline
	Japanese  	& $0.196$			& $26.791$  		& $1.079$			\\ \hline
	Turkish   	& $2.355$			& $63.199$  		& $18.449$			\\ \hline
	Yoruba    	& $1.178$			& $48.577$  		& $36.997$			\\ \hline
	Basque    	& $0.393$			& $12.365$  		& $34.838$			\\ \hline
	Swahili   	& $14.328$ 			& $85.574$  		& $0.000$			\\ \hline
	Chinese   	& $1.865$ 			& $69.382$  		& $9.225$			\\ \hline \hline
	Total     	& ? 				& ? 				& ? 				\\ \hline
	\end{tabular}
	\caption{Number of female, male and neutral pronominal genders per language in the translated sentences. The corresponding sex ratios (\# Male / \# Female) show just how much male defaults are prominent in some languages such as Chinese, with almost 30 male pronouns for each female one.}
	\label{tab:gender-by-language}
	}
\end{table}

@@

@@

\begin{figure}[H]
	\centering
	\includegraphics[width=10cm]{pictures/histogram-female-grouped}
	\caption{Histogram of the distribution of female pronouns among different occupation categories, with a distinctive peak at 0. The corresponding histogram for male pronouns (see Figure \ref{fig:histogram-male}), by constrast, peaks at 9.}
	\label{fig:histogram-female-grouped}
\end{figure}

\begin{figure}[H]
	\centering
	\includegraphics[width=10cm]{pictures/histogram-male-grouped}
	\caption{Histogram of the distribution of male pronouns among different occupation categories, with a distinctive peak at 9. In contrast with the corresponding histograms for female and gender neutral pronouns (see Figures \ref{fig:histogram-female} and \ref{fig:histogram-neutral} respectively) which exhibit inverse correlations, here one can see a tendency towards normally-distributed pronoun counts.}
	\label{fig:histogram-male-grouped}
\end{figure}

\begin{figure}[H]
	\centering
	\includegraphics[width=10cm]{pictures/histogram-neutral-grouped}
	\caption{Following the behavior observed for female pronouns, the histogram for the distribution of gender neutral pronouns among different categories shows a tendency towards an inverse correlation, with a distinctive peak at 0.}
	\label{fig:histogram-neutral-grouped}
\end{figure}

@@

\section{Conclusions}
In this paper, we have provided evidence that statistical translation tools such as Google Translate exhibit implicit gender biases and a tendency towards male defaults, which possibly stem from the real world data used to train them. As a result, we suggest that such tools can be probed to yield insights about stereotypical gender roles in our society. By translating sentences from gender neutral languages such as Hungarian and Chinese into English, we were able to collect statistics about the asymmetry between female and male pronominal genders in the translation process. Because Google Translate typically uses English as a \emph{lingua franca} to translate between other languages, our findings about gender bias in professional occupations possibly generalize to most translations from gender neutral idioms in the same tool, although we have not tested this hypothesis yet. Our results further show that although male pronouns are typical, the proportion of female pronouns can vary significantly according either to the adjectives used (\say{shy}/\say{happy} vs \say{brave}/\say{guilty}) or the category of professional occupations (artistic jobs are far more likely to be translated with female pronouns than computer-related ones). We have also shown that different languages are differently biased towards male defaults, with Hungarian exhibiting a better balance between male and female pronouns than, say, Chinese. Some languages such as Yoruba and Basque have been found to translate sentences with gender neutral pronouns very often, although this is the exception rather than the rule.

We think this work can shed further light on what we believe is a pressing ethical difficulty arising from modern statistical machine translation, and hope that it will lead to discussions about the role of AI engineers on minimizing potentially harmful effects of the pressing modern concerns of machine bias.

%\bibliography{main}
%\bibliographystyle{agsm}

%\ifx
\begin{thebibliography}{xx}

@@

\harvarditem[Bolukbasi et~al.]{Bolukbasi, Chang, Zou, Saligrama \harvardand\
  Kalai}{2016}{bolukbasi2016man}
Bolukbasi, T., Chang, K.-W., Zou, J.~Y., Saligrama, V. \harvardand\ Kalai,
  A.~T.  \harvardyearleft 2016\harvardyearright , Man is to computer programmer
  as woman is to homemaker? debiasing word embeddings, {\em in} `Advances in
  Neural Information Processing Systems', pp.~4349--4357.

@@

\harvarditem[Hajian et~al.]{Hajian, Bonchi \harvardand\
  Castillo}{2016}{hajian2016algorithmic}
Hajian, S., Bonchi, F. \harvardand\ Castillo, C.  \harvardyearleft
  2016\harvardyearright , Algorithmic bias: From discrimination discovery to
  fairness-aware data mining, {\em in} `Proceedings of the 22nd ACM SIGKDD
  international conference on knowledge discovery and data mining', ACM,
  pp.~2125--2126.

@@

\harvarditem{Kelman}{2014}{TranslateCommunity}
Kelman, S.  \harvardyearleft 2014\harvardyearright , `Translate community: Help
  us improve google translate!'.
\newblock Last visited 2018-3-12.
\newline\harvardurl{https://search.googleblog.com/2014/07/translate-community-help-us-improve.html}

\harvarditem{Kirkpatrick}{2016}{kirkpatrick2016battling}
Kirkpatrick, K.  \harvardyearleft 2016\harvardyearright , `Battling algorithmic
  bias: how do we ensure algorithms treat us fairly?', {\em Communications of
  the ACM} {\bf 59}(10),~16--17.

@@

\harvarditem{Olson}{2018}{AlgorithmGtranslateSexist2018}
Olson, P.  \harvardyearleft 2018\harvardyearright , `The algorithm that helped
  google translate become sexist'.
\newblock Last visited 2018-3-12.
\newline\harvardurl{https://www.forbes.com/sites/parmyolson/2018/02/15/the-algorithm-that-helped-google-translate-become-sexist/\#1c1122c27daa}

@@

\harvarditem{Shankland}{2017}{Gtranslate200daily2017}
Shankland, S.  \harvardyearleft 2017\harvardyearright , `Google translate now
  serves 200 million people daily'.
\newblock Last visited 2018-3-12.
\newline\harvardurl{https://www.cnet.com/news/google-translate-now-serves-200-million-people-daily/}

@@

\end{thebibliography}
%\fi

\end{document}
